---
title: "FluxCalR package"
author: "Junbin Zhao (junbinzhao1985@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FluxCalR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background
=============================
Static chamber is one of the most commonly used approach for measuring greenhouse gas flux (e.g., CO2 and CH4) from ecosystems (e.g., stem/soil respiration, CH4 emission, etc.) for its easy implementation, high accuracy and low cost. To perform the measurements, commercial infrared gas analyzers (IRGA) are usually used to record the changes of greenhouse gas concentrations within a chamber that covers an area of interest (e.g., soil surface) at a certain temporal resolution (e.g., 1s). However, to calculate the fluxes from the measured gas concentrations can be tedious and subjective considering that the original dataset usually includes large amount of unwanted records for the periods between the measurements and that the window for calculating each flux has to be defined manually. Therefore, a tool that can efficiently calculate the concentration changes into fluxes in a more objective fashion is needed.

## What does the package do?
==============================
The `FluxCalR` package directly takes the raw data exported by a gas analyzer (e.g. LGR Ultraportable Gas Analyzers) after chamber measurements and calculate fluxes of multiple measurements at once from the data file. Specifically, users can use the function `LoadLGR()` or `LoadOther()` to load the raw data file that includes the gas concentration records. Then, create a separate data frame that includes the time cues that each measurement starts and/or ends corresponding to the time in the raw data file. Based on the time cues, the fluxes of all the measurements can be calculated and exported using the function `FluxCal()`. The package has several features that can facilitate the process of flux calculation:

* The package includes the function `SelCue()` that one can use to select the time cues (start or end of each measurement) manually by clicking on an interactive graph of the entire gas concentration time series from the raw data file. The data frame returned from the function can be directly used as the argument `df_cue` of `FluxCal()` for the flux calculation. The time cue data frame can also be saved (default) and used again to get reproducible results.   

* Depending on the measurement protocol and requirement of the users, the `FluxCal()` function is flexible in the sense that 1) it takes either "Start", "End" or both "Start_End" of each measurement (argument `cue_type`) as time cues to identify each measurement; 2) it calculates "CO2", "CH4" or both "CO2_CH4" fluxes at the same time (argument `cal`). 

* The flux calculation automatically scans through a range of 1.5x (default) of the measurement window and the best linear regression (largest R2) is used to calculate each of the final fluxes. This protocol optimize the representative and reproducibility of the calculated flux. This dynamic range can be changed in the argument `ext` of the function `FluxCal()`. 

* By default (argument `check_plot`), a graph is plotted after the calculation with regression lines plotted on the CO2 and/or CH4 concentration time series for checkup purposes. This post-calculation feature will help users identify any flux calculation that is based on the unwanted section of the data. 

* The `FluxCal()` returns a data frame with the calculated flux and identifier of each measurement (e.g., plot ID, see argument `Other`) that can be integrated into users' data processing/analyzing workflow. At the same time, it also saves the data frame to a ".csv" file that can be loaded by R again later or used in other programs for non-R users. 


## Examples of using the package
================================
```{r setup}
library(FluxCalR)
```
